### 2019-02-07 outline/outline.Rmd
added outline

---
title: "Quantifying acute and chronic oak decline severity using phenotypic descriptors"
author: "Jasen Finch"
date: "07/02/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Authors:** Jasen Finch, Nathan Brown, John Draper, Sandra Denman

**Proposed journals:**

**Summary:**

**Keywords:** phenotyping, tree-decline, machine-learning, quantification

## Introduction

## Methods

## Results/Discussion

### Phenotypic descriptors

### Extended descriptors

### Decline Indexes

###

### Applications

## Conclusion