### 2019-11-14 manuscript/tables.Rmd
tables and figures temporarily rendered separate from the manuscript

---
title: Tables
output: pdf_document
---

## Tables

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r phenotypic_descriptors,cache=FALSE}
loadd(phenotypic_descriptor_info)

phenotypic_descriptor_info %>%
  kable(format = 'markdown')
```

```{r resampled_results}

```