### 2019-11-25 manuscript/tables.Rmd
added to descriptor table

---
title: Tables
output: 
  pdf_document:
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

\newpage

\begin{blandscape}

```{r phenotypic_descriptors,cache=FALSE}
loadd(phenotypic_descriptor_info)

phenotypic_descriptor_info %>%
  kable(format = 'markdown')
```

\end{elandscape}

```{r resampled_results}

```