### 2019-12-10 manuscript/plan.Rmd
plan compiles

---
title: "Plan"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\newpage
## Checks

```{r check,comment='',cache=FALSE}
manuscript_check(file = 'plan.Rmd')
```