### 2019-12-16 manuscript/supplementary.Rmd
added figure numbers

---
title: Supplementary information
output: 
  pdf_document:
    keep_tex: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r phenotypic_descriptors,cache=FALSE}
loadd(phenotypic_descriptor_info)

phenotypic_descriptor_info %>%
  select(-`Descriptor Type`,-`Data Type`) %>%
  kable(format = 'latex',
        booktabs = TRUE,
        longtable = TRUE,
        caption = 'Phenotypic descriptors collected for oak decline severity.') %>%
  kable_styling(latex_options = c('striped','repeat_header'),repeat_header_continued = TRUE) %>%
  # column_spec(1,width = '1.5cm') %>%
  column_spec(1,width = '3cm') %>%
  column_spec(2:3,width = '6cm') #%>%
  # collapse_rows(columns = 1,latex_hline = "major", valign = "middle")
```

```{r descriptor_adjustment,fig.cap='Exapmle of overall mean adjustment applied to the diameter at breast height descriptor. Error bars show ±1SE of the mean. Dashed red line shows the overall mean.'}
loadd(descriptor_adjustment_example)

descriptor_adjustment_example %>%
  wrap_plots()
```

```{r descriptor_adjustment_factors}
loadd(site_correction_factors)

site_correction_factors %>%
  mutate_if(is.double,signif,digits = 3) %>%
  kable(format = 'latex',booktabs = TRUE,caption = 'Descriptor adjustment factors at each site.') %>%
  collapse_rows(columns = 1,latex_hline = "major", valign = "middle")
```

```{r modelTuningResultsPlot,fig.cap='Random forest parameter tuning results for decline index models.'}
loadd(rf_tune_plot)

rf_tune_plot
```

```{r site_differences,fig.cap='Differences between sites in phenotypic descriptors analysed by supervised Random Forest. a) Multidimensional scaling (MDS) plot of random forest proximity values of individual trees scaled to two dimensions. Mean margin value shown. b) Relative descriptor Gini importances measures.',fig.height=6}
loadd(site_differences_mds_plot)
grid.arrange(site_differences_mds_plot$a,site_differences_mds_plot$b,nrow = 1)
```

```{r PDInumeric,fig.height=9,fig.cap='Continuous descriptors plotted against the PDI.'}
loadd(descriptor_scatter_plots)

descriptor_scatter_plots$PDI_numeric
```

```{r PDIfactor,fig.height=8,fig.cap='Categorical descriptors plotted against the PDI.'}
descriptor_scatter_plots$PDI_factor
```

```{r DAInumeric,fig.height=9,fig.cap='Continuous descriptors plotted against the DAI.'}
descriptor_scatter_plots$DAI_numeric
```

```{r DAIfactor,fig.height=8,fig.cap='Categorical descriptors plotted against the DAI.'}
descriptor_scatter_plots$DAI_factor
```