### 2020-08-17 manuscript/submission/author_contributions.Rmd
author contributions now output as word document

---
output: word_document
---

### CRediT authorship contribution statement

**Jasen Finch:** Conceptualization, Methodology, Software, Formal analysis, Data Curation, Visualization, Writing - Original Draft, Writing - review & editing.
**Nathan Brown:** Methodology, Investigation, Data Curation, Writing - Review & Editing. 
**Manfred Beckmann:** Writing - Review & Editing, Supervision.
**Sandra Denman:** Methodology, Writing - Review & Editing, Project administration, Funding acquisition.
**John Draper:** Methodology, Supervision, Project administration.